|

### 2019-12-15 article/article.Rmd
changes in formatting, a few edits to text

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare that there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does grassroots participation improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays. Our results show that the app has a null impact on the school construction outcomes. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin: 
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

```{r setup, include=FALSE, warning=FALSE}
pkgs <- c('tidyverse', 'dummies', 'kableExtra',
          'haven', 'compareGroups', 'grid', 'gridExtra',
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt', 
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg', 
          'flextable', 'officer', 'ggthemes', 
          'stargazer')

# Install required packages
installIfNot <- function(x) {
  if(x %in% rownames(installed.packages()) == FALSE) 
    install.packages(x, dependencies = T)
} 
lapply(pkgs, installIfNot)

# Load scripts
lapply(pkgs, require, character.only = T)
rm(pkgs, installIfNot)
load('tdp.RData')
```

@@

A robust accountability system is crucial for efficient public services provision [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent; @o1998horizontal]. In its standard definition, accountability is understood as the process of holding authorities responsible for their actions [@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past studies show that accountability has a positive impact on governance, as it ensures that politicians act on behalf of voters [@freire2010ngp; @moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking and corruption [@deininger2005does; @wenar2006accountability], and improves the quality of public services [@adsera2003you; @bjorkman2009power]. Recent research also suggests that accountability leads to higher economic growth because it limits state discretion in the economy and increases long-term investments in human capital [@benhabib2010economic; @suebvises2018social; @ponzetto2018social].

However, accountability systems take many forms. One promising model is that of _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project so they can evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents argue that bottom-up accountability is effective because: 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to attack corruption that directly affects themselves; 3) policy-makers are sensitive to social punishment from their own communities [@serra2011combining, 570]. In this regard, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of state officials with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

@@

We use the TDP app to conduct two experimental interventions and test its impact on five outcomes related to school completion rates and complaints to public authorities. Overall, our results show that providing information to citizens has no consistent impact on policy outcomes. In the first experiment, we find that the TDP app increased the likelihood of construction cancellation by 2 percent, but the result does not replicate. The remaining five models have null results. In our second intervention, none of the estimations reach conventional levels of statistical significance. Importantly, all coefficients are small, what suggests that even if the TDP app had a significant effect on the outcomes, its substantive impact would be virtually negligible.

The findings raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. Although studies like @bjorkman2009power, @fiala2017social, and @reinikka2005fighting report better policy outcomes after providing information to local communities, our findings are closer to @banerjee2010pitfalls, @keefer2014mass, @lieberman2014does, @bjorkman2017experimental, @olken2007monitoring, and @raffler2018weakness offering little evidence that information-based interventions lead to greater government responsiveness. Our analysis suggests that grassroots monitoring is ineffective in altering government behaviour in Brazil.

@@

Between August 2017 and February 2019, we implemented two interventions to measure the effect of the TDP app on five school construction outcomes plus a placebo test. The outcomes are: 1) a placebo outcome indicating the percentage of the project completed before the impact evaluation started; 2) the percentage of the project completed by the end of the interventions; 3) the difference between the percentage reported as completed before and after the interventions; 4) the number of finished constructions; 5) the number of cancelled constructions; 6) the number of schools where construction companies updated the conclusion dates. In terms of expectations, Table \ref{tab:outcomes} depicts the expected effects for each of the studied outcomes.

@@

# Load data 
load('tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups', 
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt', 
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg', 
          'flextable', 'officer', 'ggthemes')
invisible(lapply(pkgs, require, character.only = T))

## Map GA intervention 1
mapbr <- get_brmap('City')
gaaux <- filter(ga, phase1 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase1 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 2
gaaux$Downloads <- factor(gaaux$downl, labels = c('No Downloads', 'Downloads', 'Not in App'))
mapbr <- join_data(mapbr, gaaux, by = 'City')
pMapDownloadsInt1 <- ggplot() + 
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values=c( "#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
pMapDownloadsInt1

# Frequency downloads
tab <- filter(ga, phase1 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c('Date', 'Frequency')
pFreqUserDownloadsInt1 <- ggplot(tab, aes(x=Date, y=Frequency, group=1)) +
  geom_line() + geom_point()
pFreqUserDownloadsInt1
```

\newpage 
The second intervention is similar to intervention 1 in all but three characteristics. First, the TDP app was available for both Android and iOS devices. Second, we randomised the intervention at the school level, with `r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and `r as.numeric(table(impact_evaluation_phase2$treat)[2])` treatment units. We used blocked randomisation stratified by Brazilian states, school construction status (under construction, stopped, unfinished), and whether the municipality spent more on school construction that the distribution median. Finally, the intervention period lasted from August 2018 to February 2019.

@@

```{r man2, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=3.5, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation2}Manipulation checks for intervention 2. The graphs display the geographical and temporal variation of TDP app downloads from August 2018 to February 2019.'}
# Load data 
load('tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups', 
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt', 
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg', 
          'flextable', 'officer', 'ggthemes')
invisible(lapply(pkgs, require, character.only = T))

## Manipulation Stats

## Map manipulation test -- Intervention 2
mapbr <- get_brmap('City')
gaaux <- filter(ga, phase2 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase2 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 2
gaaux$Downloads <- factor(gaaux$downl, labels = c('No Downloads', 'Downloads',
                                                  'Not in App'))
mapbr <- join_data(mapbr, gaaux, by = 'City')
p1 <- ggplot() + 
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values=c( "#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
p1

# Downloads over time -- Intervention 2:
# Phase 2 -- manipulation
tab <- filter(ga, phase2 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c('Date', 'Frequency')
p2 <- ggplot(tab, aes(x=Date, y=Frequency, group=1)) +
  geom_line() + geom_point()
p2
```

@@

\begin{equation} 
Y_i \ = \ \alpha + \beta T_i + \gamma X_i + \theta Z_i + \varepsilon_i 
\end{equation}

where $i$ indexes the experiment units. $Y_i$ is one of the six outcomes described above, $\alpha$ is the intercept, $\beta$ denotes the average treatment effect, and $T_i$ is a binary treatment indicator. $\gamma$ is a vector of fixed effects, $X_i$ is a matrix of Brazilian states' fixed effects, $\theta$ is a vector of controls and $Z_i$ an array of controls for the case $i$. The error term is denoted by $\varepsilon_i$. We cluster the standard errors at the municipality level as mayors are responsible for school investment decisions in Brazil.

@@

```{r int1, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
load('tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups', 
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt', 
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg', 
          'flextable', 'officer', 'ggthemes')
invisible(lapply(pkgs, require, character.only = T))

# Auxiliary function: calculate coefficients
calcATEs <- function(outcs, treats, ctrls, fes, cls, dat, wts = NULL, ...) {
  ctrls_mod <- paste(ctrls, collapse = '+')
  fes_mod <- paste(fes, collapse = '+')
  moddescr <- data.frame()
  mods <- list()
  dataux <- dat
  k <- 1
  for (i in treats) {
    for (j in outcs) {
      if(is.null(wts)) {
        dat <- dataux %>%
          select(one_of(c(j, treats, ctrls, fes, cls))) %>%
          na.omit()
        } else {
          dat <- dataux %>%
            select(one_of(c(j, treats, ctrls, fes, cls, wts))) %>%
            na.omit()
          wts <- dat[,wts]
        }
      # No controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(i,paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wts)
      moddescr <- rbind(moddescr, 
                        data.frame(outc = j, 
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = F))
      print(k); k = k+1
      # Controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wts)
      moddescr <- rbind(moddescr, 
                        data.frame(outc = j, 
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = F))
      print(k); k = k+1
      # No controls and fes
      mods[[k]] <- felm(
        as.formula(paste(c(j,paste(paste(c(i), collapse = '+'),
                                   paste('',fes_mod,0,cls, sep = '|'), 
                                   sep = '')),
                         collapse = '~')), data = dat, weights = wts)
      moddescr <- rbind(moddescr, 
                        data.frame(outc = j, 
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = T))
      print(k); k = k+1
      # Controls and fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',fes_mod,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wts)
      moddescr <- rbind(moddescr, 
                        data.frame(outc = j, 
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = T))
      print(k); k = k+1
    }
  }
  return(list(moddescr, mods))
}

@@

## ATEs
outcs <- c('percExecBegin', 'percExecEnd',
           'deltaPercExec', 'isFinished',
           'isCancelled', 'updDate')
treats <- c('treat')
ctrls <- c('logPop2015', 'ideb_ai_2015','ideb_af_2015',
           'log_poorFam2010', 'log_totTransf2016')
fes <- c('state')
cls <- c('municipality')

resInt1 <- calcATEs(outcs, treats, ctrls, fes, cls, dat = impact_evaluation_phase1)

# Results intervention 1
stargazer(resInt1[[2]][[4]],resInt1[[2]][[8]],resInt1[[2]][[12]],
          resInt1[[2]][[16]],resInt1[[2]][[20]],resInt1[[2]][[24]],
          title = '\\label{tab:results1} Impact Evaluation -- Intervention 1', 
          keep='treatTreatment', 
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = 'Cluster-robust SEs at the municipality level.',
          add.lines = list(c("Controls", "Yes", "Yes", "Yes", 
                             "Yes", "Yes", "Yes"),
            c("State fixed effects", "Yes", "Yes", "Yes",
              "Yes", "Yes", "Yes")),
          notes.append = T, header=FALSE)
```

@@

We find that the app only has a small effect on cancellation rates. The TDP application increases the likelihood of cancelling the construction in `r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result goes in the opposite direction of our theoretical expectations, the finding is largely inconsistent, and it does not replicate in the second experiment. All other coefficients are not statistically significant at conventional levels. On the one hand, the results indicate that our placebo outcome, the percentage of the invested executed before the intervention, indeed behaves as predicted. On the other hand, we expected the five remaining outcomes to improve after the introduction of the app. The literature on bottom-up accountability argues that more information about the shortcomings of public service provision will mobilise citizens who, in turn, would pressure state agents to produce better social policies. Our results do not lend support to that hypothesis.

@@

```{r int2, echo=FALSE, warning=FALSE, message=FALSE, results='hide',cache=TRUE}
load('tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups', 
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt', 
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg', 
          'flextable', 'officer', 'ggthemes', 
          'stargazer')
invisible(lapply(pkgs, require, character.only = T))

>>>>

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author:
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare that there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does ~~grassroots participation~~ **local oversight** improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin:
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

```{r setup, include=FALSE, warning=FALSE}
pkgs <- c("tidyverse", "dummies", "kableExtra",
          "haven", "compareGroups", "grid", "gridExtra",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes",
          "stargazer")

# Install required packages
install_if_not <- function(x) {
  if (x %in% rownames(installed.packages()) == FALSE)
    install.packages(x, dependencies = T)
}
lapply(pkgs, install_if_not)

# Load scripts
lapply(pkgs, require, character.only = T)
rm(pkgs, install_if_not)
load("tdp.RData")
```

@@

A robust accountability system is crucial for efficient public services
provision [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent;
@o1998horizontal]. In its standard definition, accountability is understood as
the process of holding authorities responsible for their actions
[@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past
studies show that accountability has a positive impact on governance; ~~as~~ it
ensures that politicians act on behalf of voters [@freire2010ngp;
@moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking
and corruption [@deininger2005does; @wenar2006accountability], and improves the
quality of public services [@adsera2003you; @bjorkman2009power]. Recent
research also suggests that accountability leads to higher economic growth
because it limits state discretion in the economy and increases long-term
investments in human capital [@benhabib2010economic; @suebvises2018social;
@ponzetto2018social].

However, accountability systems **may** take many forms. One promising model is
that of _bottom-up monitoring_, in which citizens receive information about the
shortcomings of a given project so they can evaluate and pressure
underperforming public officials [@kosack2014does; @molina2016community;
@raffler2018weakness]. Proponents argue that bottom-up accountability is
effective because: 1) constituents have first-hand information about the
outcomes of local policies; 2) citizens have incentives to attack corruption
that directly affects themselves; 3) policy-makers are sensitive to social
punishment from their own communities [@serra2011combining, 570]. In this
regard, bottom-up accountability offers a potential solution to the
principal-agent dilemma in public service by aligning the interests of state
officials with those of the constituency they serve [@barro1973control;
@raffler2018weakness, 2].

@@

We use the TDP app to conduct two experimental interventions and test ~~its
impact~~ **the impact of citizen oversight** on five outcomes related to school
completion rates and complaints to public authorities. Overall, our results
show that providing information to citizens has no consistent impact on policy
outcomes. In the first experiment, we find that the TDP app increased the
likelihood of construction cancellation by 2 percent, but the result does not
replicate. The remaining five models have null results. In our second
intervention, none of the estimations reach conventional levels of statistical
significance. Importantly, all coefficients are small, what suggests that even
if the TDP app had a significant effect on the outcomes, its substantive impact
would be ~~virtually~~ negligible.

The findings raise questions about the ability of citizens to hold
representatives accountable using bottom-up ~~strategies~~ **monitoring**.
Although studies like @bjorkman2009power, @fiala2017social, and
@reinikka2005fighting report better policy outcomes after providing information
to local communities, our findings are closer to @banerjee2010pitfalls,
@keefer2014mass, @lieberman2014does, @bjorkman2017experimental,
@olken2007monitoring, and @raffler2018weakness, offering little evidence that
information-based interventions lead to greater government responsiveness. Our
analysis suggests that grassroots monitoring is ineffective in altering
government behaviour in Brazil.

@@

Between August 2017 and February 2019, we implemented two interventions to
measure the effect of the TDP app on five school construction outcomes plus
a placebo test. The outcomes are: 1) a placebo outcome indicating the
percentage of the project completed before the impact evaluation started; 2)
the percentage of the project completed by the end of the interventions; 3) the
difference between the percentage reported as completed before and after the
interventions; 4) the number of finished constructions; 5) the number of
cancelled constructions; 6) the number of schools where construction companies
updated the conclusion dates. In terms of expectations, Table
\ref{tab:outcomes} depicts the expected effects for each of the studied
outcomes[^more-details].

[^more-details]: **Please refer to the Online Appendix for further details on
the treatment implementation and the coding of the outcome variables.**

@@

# Load data
load("tdp.RData")

# Load packages
pkgs <- c("tidyverse", "dummies", "ggpubr",
          "haven", "compareGroups",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes")
invisible(lapply(pkgs, require, character.only = T))

## Map GA intervention 1
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase1 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$city <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase1 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl,
													labels = c("No Downloads", "Downloads", "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
plot_downloads_int1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
plot_downloads_int1

# Frequency downloads
tab <- filter(ga, phase1 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c("Date", "Frequency")
plot_freq_users_int1 <- ggplot(tab, aes(x = Date, y = Frequency, group = 1)) +
  geom_line() + geom_point()
plot_freq_users_int1
```

\newpage
The second intervention is similar to intervention 1 in all but three
characteristics. First, the TDP app was **then** available for both Android and
iOS devices. Second, we randomised the intervention at the school level, with
`r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and 
`r as.numeric(table(impact_evaluation_phase2$treat)[2])` treatment units. We
used blocked randomisation stratified by Brazilian states, school construction
status (under construction, stopped, unfinished), and whether the municipality
spent more on school construction that the distribution median. Finally, the
intervention period lasted from August 2018 to February 2019.

@@

```{r man2, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=3.5, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation2}Manipulation checks for intervention 2. The graphs display the geographical and temporal variation of TDP app downloads from August 2018 to February 2019.'}
# Load data
load("tdp.RData")

# Load packages
pkgs <- c("tidyverse", "dummies", "ggpubr",
          "haven", "compareGroups",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes")
invisible(lapply(pkgs, require, character.only = T))

## Manipulation Statistics

## Map manipulation test -- Intervention 2
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase2 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase2 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl, labels = c("No Downloads", "Downloads",
                                                  "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
p1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
p1

# Downloads over time -- Intervention 2:
# Phase 2 -- manipulation
tab <- filter(ga, phase2 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c("Date", "Frequency")
p2 <- ggplot(tab, aes(x = Date, y = Frequency, group = 1)) +
  geom_line() + geom_point()
p2
```

**(Short paragraph on whether the app has worked)**

@@

\begin{equation}
Y_i = \alpha + \beta T_i + \gamma X_i + \theta Z_i + \varepsilon_i
\end{equation}

where $i$ indexes the experiment units. $Y_i$ is one of the six outcomes
described above, $\alpha$ is the intercept, $\beta$ denotes the average
treatment effect, and $T_i$ is a binary treatment indicator. $\gamma$ is
a vector of fixed effects, $X_i$ is a matrix of Brazilian states' fixed
effects, $\theta$ is a vector of controls, and $Z_i$ an array of controls for
the case $i$. The error term is denoted by $\varepsilon_i$. We cluster the
standard errors at the municipality level as mayors are responsible for school
investment decisions in Brazil.

@@

```{r int1, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
load("tdp.RData")

# Load packages
pkgs <- c("tidyverse", "dummies", "ggpubr",
          "haven", "compareGroups",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes")
invisible(lapply(pkgs, require, character.only = T))

# Auxiliary function: calculate coefficients
calc_ates <- function(outcs, treats, ctrls, fes, cls, dat, wts = NULL, ...) {
  ctrls_mod <- paste(ctrls, collapse = "+")
  fes_mod <- paste(fes, collapse = "+")
  moddescr <- data.frame()
  mods <- list()
  dataux <- dat
  k <- 1
  for (i in treats) {
    for (j in outcs) {
      if (is.null(wts)) {
        dat <- dataux %>%
          select(one_of(c(j, treats, ctrls, fes, cls))) %>%
          na.omit()
        } else {
          dat <- dataux %>%
            select(one_of(c(j, treats, ctrls, fes, cls, wts))) %>%
            na.omit()
          wts <- dat[, wts]
        }
      # No controls and no fixed effects
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(i,paste("", 0, 0, cls, sep = "|"), sep = "")),
              collapse = "~")), data = dat, weights = wts)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = F))
      print(k); k = k + 1
      # Controls and no fixed effects
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i, ctrls_mod), collapse = "+"),
                        paste("", 0, 0, cls, sep = "|"), sep = "")),
              collapse = "~")), data = dat, weights = wts)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = F))
      print(k); k = k + 1
      # No controls and fixed effects
      mods[[k]] <- felm(
        as.formula(paste(c(j, paste(paste(c(i), collapse = "+"),
                                   paste("", fes_mod, 0, cls, sep = "|"),
                                   sep = "")),
                         collapse = "~")), data = dat, weights = wts)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = T))
      print(k); k = k + 1
      # Controls and fixed effects
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i, ctrls_mod), collapse = "+"),
                        paste("", fes_mod, 0, cls, sep = "|"), sep = "")),
              collapse = "~")), data = dat, weights = wts)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = T))
      print(k); k = k + 1
    }
  }
  return(list(moddescr, mods))
}

@@

## ATEs
outcs <- c("percExecBegin", "percExecEnd",
           "deltaPercExec", "isFinished",
           "isCancelled", "updDate")
treats <- c("treat")
ctrls <- c("logPop2015", "ideb_ai_2015","ideb_af_2015",
           "log_poorFam2010", "log_totTransf2016")
fes <- c("state")
cls <- c("municipality")

resInt1 <- calc_ates(outcs, treats, ctrls, fes, cls,
										 dat = impact_evaluation_phase1)

# Results intervention 1
stargazer(resInt1[[2]][[4]],resInt1[[2]][[8]],resInt1[[2]][[12]],
          resInt1[[2]][[16]],resInt1[[2]][[20]],resInt1[[2]][[24]],
          title = "\\label{tab:results1} Impact Evaluation -- Intervention 1",
          keep="treatTreatment",
          type = "latex",
          covariate.labels = "ATE",
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include = F, column.sep.width = "-2pt",
          omit.stat=c("adj.rsq", "ser"), align = F, digits = 2,
          notes = "Cluster-robust SEs at the municipality level.",
          add.lines = list(c("Controls", "Yes", "Yes", "Yes",
                             "Yes", "Yes", "Yes"),
            c("State fixed effects", "Yes", "Yes", "Yes",
              "Yes", "Yes", "Yes")),
          notes.append = T, header = F)
```

@@

We find that the app only has a small effect on cancellation rates. The TDP
application increases the likelihood of cancelling the construction in 
`r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result
goes in the opposite direction of our theoretical expectations, the finding is
inconsistent and does not replicate in the second experiment. All other
coefficients are not statistically significant at conventional levels. On the
one hand, the results indicate that our placebo outcome, the percentage of the
invested executed before the intervention, indeed behaves as predicted. On the
other hand, we expected the five remaining outcomes to improve after the
introduction of the app. The literature on bottom-up accountability argues that
delivering more information about the shortcomings of public services provision
will ~~mobilise citizens who, in turn, would pressure state agents to produce
better social policies~~ **put citizens in a position where they can monitor
state agents and improve provider behaviour** [@raffler2018weakness, 2] . Our
results do not lend support to that hypothesis.

@@

```{r int2, echo=FALSE, warning=FALSE, message=FALSE, results='hide',cache=TRUE}
load("tdp.RData")

# Load packages
pkgs <- c("tidyverse", "dummies", "ggpubr",
          "haven", "compareGroups",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes",
          "stargazer")
invisible(lapply(pkgs, require, character.only = T))
